---
title: "Gender Equality in Norwegian Municipalities"
author: "Kieran Healy"
date: "4/7/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rgdal)

library(tidyverse)

theme_set(theme_minimal())

theme_normap <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) %+replace%
    theme(axis.line=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          axis.title=element_blank(),
          panel.background=element_blank(),
          panel.border=element_blank(),
          panel.grid=element_blank(),
          panel.spacing=unit(0, "lines"),
          plot.background=element_blank(),
          legend.justification = c(0,0),
          legend.position = c(0.6,0.4)
          )
}


## Make a "figures" subdirectory if one doesn't exist
ifelse(!dir.exists(file.path("figures")),
       dir.create(file.path("figures")),
       FALSE)

```

# Load the Map Data

Annoyingly, these geojson files include Norway's sea boundary, and so
the coastline resembles a large blob. But we'll continue for purposes
of illustration.

First we convert the data from geojson to something we can work with, and set the projection.

```{r }

norway_komm <- readOGR(dsn = "geojson/kommuner.geojson",
                       layer = "OGRGeoJSON")

norway_komm <- spTransform(norway_komm,
                           CRS("+proj=laea +lat_0=59.91 +lon_0=10.75 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))

```

```{r}
dim(norway_komm)
```

The `noraway_komm` object appears to be a small data frame, but in fact it is a complext spatial object with lots of components (as a call to `str(norway_komm)` will confirm). The 434 features it has are the kommuner units. 

What we need to do is "reach inside" the spatial data object and create an id variable for the kommuner units that we will use to match on shortly. 

```{r}
norway_komm@data$id <- rownames(norway_komm@data)

```

The `@data$id` and `@data` instructions are ways of referring to components of the `norway_komm` object.

Next we "fortify" the `norway_komm` spatial object, which in essence means turning it into a data frame of the sort we are used to working with while mapping things.

```{r}
kommuner_map <- fortify(norway_komm)

```

Next, we need to match the `id` variables and extract the kommune names. We create an index variable, `ind`, and get the names of the kommuner from `norway_komm@data$navn`. We use our index variable to make sure we get our match right.

```{r}
ind <- match(kommuner_map$id, norway_komm@data$id)

kommuner_map$kommune <- norway_komm@data$navn[ind]
```

Finally, for this stage, we convert the kommuner names from a factor to a character type, which will make them easier to deal with.

```{r}
kommuner_map <- kommuner_map %>% mutate_if(is.factor, as.character)
```

Now we are in a position to draw a map of the kommuner. Note that we are going to save it directly to disk as a PDF. Drawing maps to the screen can often be inefficient and extremely time-consuming.

```{r}
p <- ggplot(data = kommuner_map,
            mapping = aes(x = long,
                          y = lat,
                          group = group))


p1 <- p + geom_map(data = kommuner_map,
             map = kommuner_map,
             mapping = aes(map_id = id,
                           group = group),
             color = "gray80",
             fill = "white")

pdf(file="figures/kommuner.pdf", width = 6, height = 10)
print(p1)
dev.off()

```


# Merge in some data

Statistics Norway generously provides some data to us. We join it to our map data as we did in the US case, except here the key variable is `kommune`. Then we can draw some choropleth maps. Sorry about the coastline issue!

```{r }

data <- read_csv("data/norway-gender.csv")

kommuner_map <- left_join(kommuner_map, data, by = "kommune")

p <- ggplot(data = kommuner_map,
            mapping = aes(x = long,
                          y = lat,
                          group = group))

p1 <- p + geom_map(data = kommuner_map,
                   map = kommuner_map,
                   mapping = aes(map_id = id,
                                group = group,
                                fill = cut(income_mwr, 5, dig.lab = 2)),
                   color = "grey40",
                   size = 0.1)

p2 <- p1 + coord_equal() + theme_normap() +
    labs(fill = "Ratio of Women to Men's'\nGross Income, by Kommune") + 
    scale_fill_brewer(palette = "Oranges")

pdf(file="figures/kommuner-wages.pdf", width = 6, height = 10)
print(p2)
dev.off()





```
